What We Eval in the Shadows

### 2021-09-03 paper/main.tex
Fix typos and grammar in Section 8

The \eval functions are used widely and in varied ways in R. The function is an
essential tool for language implementers. Our analysis observed that the base
libraries heavily depend on \eval and any R program will end up calling it
through the core of the language. Independently developed libraries often use
\eval in subtle ways, mostly to perform macro-programming tasks. Our review of
code written by less experienced users suggest that \eval is exceedingly rarely
used. So, it is fair to conclude that a majority of R programmers never have to
write a call to \eval, but that the code they write would not run without it.

\Eval is thus a challenge for automated program understanding. Program analysis
and transformation tools or compilers must assume the worst when faced with code
that calls this function. We have observed all sorts of side-effects with
various degree of visibility.

While our results are not encouraging in general, we have observed many cases
where \eval is used in a disciplined and predictable ways. While in the general
case \eval is hell, there are many cases where it is just a function. It does
not appear that a general purpose replacement for \eval is possible, in future
work we hope to focus on special cases and propose tools that target particular
subsets of call sites with some common properties.

>>>>

The \eval function is used widely, and in varied ways, in R. The function is an
essential tool for language implementers. Our analysis observed that the base
libraries heavily depend on \eval, and any R program will end up calling it
through the core of the language. Independently developed libraries often use
\eval in subtle ways, mostly to perform macro-programming tasks. Our review of
code written by less experienced users suggests that \eval is exceedingly rarely
used. So, it is fair to conclude that a majority of R programmers never have to
write a call to \eval, but that the code they write would not run without it.

\Eval is thus a challenge for automated program understanding. Program analysis
and transformation tools or compilers must assume the worst when faced with code
that calls this function. We have observed all sorts of side-effects with
various degrees of visibility.

While our results are not encouraging in general, we have observed many cases
where \eval is used in disciplined and predictable ways. While in the general
case \eval is hell, there are many cases where it is just a function. It does
not appear that a general-purpose replacement for \eval is possible; in future
work, we hope to focus on special cases and propose tools that target particular
subsets of call sites with some common properties.
